API design for machine learning software: experiences from the scikit-learn project

### 2013-06-25 paper.tex
minor rewords

The \sklearn project\footnote{\url{http://scikit-learn.org}}
\citep{pedregosa2011} is an open source machine learning
library for the Python programming language. The ambition of the project is to
provide efficient and well-established machine learning tools within a
programming environment that is accessible to non-machine learning experts and
reusable in various scientific areas. The project is not a novel domain-specific
language, but a library that provides machine learning idioms to a general-
purpose high-level language. Among other things, it includes classical learning
algorithms; model evalutation, selection and evaluation tools; and
preprocessing procedures. The project is
distributed under the simplified BSD license, encouraging its use in both
academic and commercial settings.

@@

Another package that deserves mention is the Gensim topic modeling toolkit
\citep{rehurek2010gensim}.
Although its stated goal is not machine learning,
it exemplifies API design for scalable processing of ``big data''
that takes a different approach than comparable models in \sklearn.
Gensim's way of dealing with large datasets is to use algorithms
that have $O(1)$ space complexity and can be updated online.
The API is designed around the Python concept of an \textit{iterable}
(supported in the language by a restricted form of co-routines called
\textit{generators}).
While the text vectorization part of \sklearn
also uses iterables to some extent,
the emphasis is still on (mini-)batch processing,
even in the stateless, O(1) memory vectorizers
that implement the hashing trick of \citet{weinberger2009}.
This way of processing, as argued earlier in Section~\ref{sec:arrays},
reduces various forms of overhead
and makes effective use of the abundance of memory in modern computers.
We make no attempt to hide this batch-oriented processing from the user,
allowing control over the amount of memory actually dedicated
to \sklearn algorithms.

>>>>

The \sklearn project\footnote{\url{http://scikit-learn.org}}
\citep{pedregosa2011} is an open source machine learning
library for the Python programming language. The ambition of the project is to
provide efficient and well-established machine learning tools within a
programming environment that is accessible to non-machine learning experts and
reusable in various scientific areas. The project is not a novel domain-specific
language, but a library that provides machine learning idioms to a general-
purpose high-level language. Among other things, it includes classical learning
algorithms, model evalutation and selection tools, as well as
preprocessing procedures. The project is
distributed under the simplified BSD license, encouraging its use in both
academic and commercial settings.

@@

Another package that deserves mention is the Gensim topic modeling toolkit
\citep{rehurek2010gensim}.
Although its stated goal is not machine learning,
it exemplifies API design for scalable processing of ``big data''
with an approach different from the comparable models in \sklearn.
Gensim's way of dealing with large datasets is to use algorithms
that have $O(1)$ space complexity and can be updated online.
The API is designed around the Python concept of an \textit{iterable}
(supported in the language by a restricted form of co-routines called
\textit{generators}).
While the text vectorization part of \sklearn
also uses iterables to some extent,
the emphasis is still on (mini-)batch processing,
even in the stateless, O(1) memory vectorizers
that implement the hashing trick of \citet{weinberger2009}.
This way of processing, as argued earlier in Section~\ref{sec:arrays},
reduces various forms of overhead
and makes effective use of the abundance of memory in modern computers.
We make no attempt to hide this batch-oriented processing from the user,
allowing control over the amount of memory actually dedicated
to \sklearn algorithms.
